---
title: "Scatterplots"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
filters:
  - webr
webr:
  packages:
    - ggplot2
    - dplyr
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
set.seed(1234)

source(here::here("R", "helpers.R"))
```

### Review 1: `geom_point()` {.no-hide}

In [Visualization Basics](/basics/01-visualization-basics/), you learned how to make a scatterplot with `geom_point()`.

The code below summarizes the mpg data set and begins to plot the results. Finish the plot with `geom_point()`. Put `mean_cty` on the $x$ axis and `mean_hwy` on the $y$ axis.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot()


```

## {{< fa circle-check >}}&ensp;Solution

```r
mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() +
    geom_point(mapping = aes(x = mean_cty, y = mean_hwy))
```

:::

###

Good job! It can be tricky to remember when to use `|>` and when to use `+`. Use `|>` to add one complete step to a pipe of code. Use `+` to add one more line to a {ggplot2} call.

### `geom_text()` and `geom_label()`

`geom_text()` and `geom_label()` create scatterplots that use words instead of points to display data. Each requires the extra aesthetic `label`, which you should map to a variable that contains text to display for each observation.

Convert the plot below from `geom_point()` to `geom_text()` and map the `label` aesthetic to the `class` variable. When you are finished convert the code to `geom_label()` and rerun the plot. Can you spot the difference?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() +
    geom_point(mapping = aes(x = mean_cty, y = mean_hwy))


```

## {{< fa circle-check >}}&ensp;Solution

```r
mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() +
    geom_text(mapping = aes(x = mean_cty, y = mean_hwy, label = class))

mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() +
    geom_label(mapping = aes(x = mean_cty, y = mean_hwy, label = class))
```

:::

###

Good job! `geom_text()` replaces each point with a piece of text supplied by the label aesthetic. `geom_label()` replaces each point with a textbox. Notice that some pieces of text overlap each other, and others run off the page. We'll soon look at a way to fix this.

### `geom_smooth()`

In [Visualization Basics](/basics/01-visualization-basics/), you met `geom_smooth()`, which provides a summarized version of a scatterplot.

`geom_smooth()` uses a model to fit a smoothed line to the data and then visualizes the results. By default, `geom_smooth()` fits a loess smooth to data sets with less than 1,000 observations, and a generalized additive model to data sets with more than 1,000 observations. 

```{r echo=FALSE, out.width="100%", message=FALSE, warning=FALSE}
#| layout-ncol: 2
mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() +
    geom_point(mapping = aes(x = mean_cty, y = mean_hwy)) +
    labs(title = "geom_point()") +
    ylim(16, 30)

mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() +
    geom_smooth(mapping = aes(x = mean_cty, y = mean_hwy), se = FALSE) +
    labs(title = "geom_smooth()") +
    ylim(16, 30)
```

### `method`

You can use the `method` parameter of `geom_smooth()` to fit and display other types of model lines. To do this, pass `method` the name of an R modeling function for `geom_smooth()` to use, such as `"lm"` (for linear models) or `"glm"` (for generalized linear models).

In the code below, use `geom_smooth()` to draw the linear model line that fits the data.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() 


```

## {{< fa circle-check >}}&ensp;Solution

```r
mpg |> 
  group_by(class) |> 
  summarize(mean_cty = mean(cty), mean_hwy = mean(hwy)) |> 
  ggplot() +
    geom_smooth(mapping = aes(x = mean_cty, y = mean_hwy), method = "lm")
```

:::

###

Good job! Now let's look at a way to make `geom_smooth()` much more useful.

##

```{r}
#| echo: false
#| results: asis
create_buttons("07-layers.html")
```
