---
title: "Continuous variables"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```

**Histograms** are perhaps the most popular way to visualize continuous distributions. Here we will look at them and their derivatives. You will learn how to:

* Make and interpret histograms
* Adjust the **binwidth** of a histogram to reveal new information
* Use geoms that are similar to histograms, such as __dotplots__, __frequency polygons__, and __densities__


**Boxplots** display the relationship between a continuous variable and a categorical variable. **Count** plots display the relationship between two categorical variables. In this tutorial, you will learn how to use both. You can learn how to:

* Make and interpret boxplots
* Rotate boxplots by flipping the coordinate system of your plot
* Use *violin* plots and *dotplots*, two geoms that are similar to boxplots
* Make and interpret count plots



A **scatterplot** displays the relationship between two continuous variables. Scatterplots are one of the most common types of graphs---in fact, you've met scatterplots already in [Visualization Basics](/01-basics/01-visualization-basics/). You can learn how to:

* Make new types of scatterplots with `geom_text()` and `geom_jitter()`
* Add multiple **layers** of geoms to a plot
* Enhance scatterplots with `geom_smooth()`, `geom_rug()`, and `geom_repel()`
* Change the **coordinate system** of a plot


A **line graph** displays a functional relationship between two continuous variables. A **map**  displays spatial data. The two may seem different, but they are made in similar ways. This tutorial will examine them both. You can learn how to:

* Make new types of line plots with `geom_step()`, `geom_area()`, `geom_path()`, and `geom_polygon()`
* Avoid "whipsawing" with the group aesthetic
* Find and plot map data with `geom_sf()`
* Transform a coordinate system into a map projection with `coord_sf()`


The tutorials were adapted from [_R for Data Science_](https://r4ds.had.co.nz/) by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at [shop.oreilly.com](http://shop.oreilly.com/product/0636920034407.do).

The tutorial uses the {ggplot2} and {dplyr} packages, which have been pre-loaded for your convenience.


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-histograms.html")
```
