---
title: "`group_by()` and `summarize()`"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
filters:
  - webr
webr:
  packages:
    - babynames
    - dplyr
    - ggplot2
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(babynames)

tops <- babynames::babynames |> 
  group_by(name, sex) |> 
  summarise(total = sum(n)) |> 
  ungroup() |> 
  top_n(10, total)
  
top_10 <- babynames::babynames |> 
  semi_join(tops, by = c("name", "sex"))

source(here::here("R", "helpers.R"))
```

```{webr-r}
#| context: setup
tops <- babynames::babynames |> 
  group_by(name, sex) |> 
  summarise(total = sum(n)) |> 
  ungroup() |> 
  top_n(10, total)
  
top_10 <- babynames::babynames |> 
  semi_join(tops, by = c("name", "sex"))
```

`group_by()` takes a data frame and then the names of one or more columns in the data frame. It returns a copy of the data frame that has been "grouped" into sets of rows that share identical combinations of values in the specified columns. 

### `group_by()` in action

For example, the result below is grouped into rows that have the same combination of `year` and `sex` values: boys in 1880 are treated as one group, girls in 1880 as another group and so on.

```{r}
babynames |>
  group_by(year, sex)
```

### Using `group_by()`

By itself, `group_by()` doesn't do much. It assigns grouping criteria that is stored as metadata alongside the original data set. If your dataset is a tibble, as above, R will tell you that the data is grouped at the top of the tibble display. In all other aspects, the data looks the same.

However, when you apply a {dplyr} function like `summarize()` to grouped data, dplyr will execute the function in a groupwise manner. Instead of computing a single summary for the entire data set, {dplyr} will compute individual summaries for each group and return them as a single data frame. The data frame will contain the summary columns as well as the columns in the grouping criteria, which makes the result decipherable

Watch these animations to see what happens with one group:

::: {.panel-tabset}
## `group_by(cat1)`

```{=html}
<video controls width="100%" style="display: block; margin: auto;">
  <source src="video/grp-summarize-01.mp4" type="video/mp4">
</video>
```

## `group_by(cat2)`

```{=html}
<video controls width="100%" style="display: block; margin: auto;">
  <source src="video/grp-summarize-02.mp4" type="video/mp4">
</video>
```
:::

Let's see what happens when we group with two columns:

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
babynames |>
  group_by(year, sex) |> 
  summarize(total = sum(n))


```

:::

To understand exactly what `group_by()` is doing, remove the line `group_by(year, sex) |>` from the code above and rerun it. How do the results change?

This animation should help with the intuition of grouping by two columns:

```{=html}
<video controls width="100%" style="display: block; margin: auto;">
  <source src="video/grp-summarize-03.mp4" type="video/mp4">
</video>
```

### Ungrouping 1

If you apply `summarize()` to grouped data, `summarize()` will return data that is grouped in a similar, but not identical fashion. `summarize()` will remove the last variable in the grouping criteria, which creates a data frame that is grouped at a higher level. For example, this `summarize()` statement receives a data frame that is grouped by year and sex, but it returns a data frame that is grouped only by year.

```{r grouping-2}
babynames |>
  group_by(year, sex) |> 
  summarize(total = sum(n))
```

### Ungrouping 2

If only one grouping variable is left in the grouping criteria, `summarize()` will return an ungrouped data set. This feature let's you progressively "unwrap" a grouped data set:

If we add another `summarize()` to our pipe, 

1.  our data set will first be grouped by year and sex.  
1. _Then_ it will be summarized into a data set grouped by year (i.e. the result above)  
1. _Then_ be summarized into a final data set that is not grouped.

```{r}
babynames |>
  group_by(year, sex) |> 
  summarize(total = sum(n)) |> 
  summarize(total = sum(total))
```

### Ungrouping 3

If you wish to manually remove the grouping criteria from a data set, you can do so with `ungroup()`.

```{r grouping-3}
babynames |>
  group_by(year, sex) |> 
  ungroup()
```

### Ungrouping 3

And, you can override the current grouping information with a new call to `group_by()`.

```{r grouping-4}
babynames |>
  group_by(year, sex) |> 
  group_by(name)
```

That's it. Between `group_by()`, `summarize()`, and `ungroup()`, you have a toolkit for taking groupwise summaries of your data at various levels of grouping.

### The most popular names by total children

You now know enough to calculate the most popular names by total children (it may take some strategizing, but you can do it!). 

In the code chunk below, use `group_by()`, `summarize()`, and `arrange()` to display the ten most popular names. Compute popularity as the total number of children of a single gender given a name. In other words, the total number of boys named "Kelly" should be computed separately from the total number of girls named "Kelly".

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
babynames |>
  group_by(name, sex) |> 
  summarize(total = sum(n)) |> 
  arrange(desc(total))
```

:::


### The history of the most popular names by total children

Let's examine how the popularity of popular names has changed over time. To help us, I've made `top_10`, which is a version of `babynames` that is trimmed down to just the ten most popular names from above.

```{r}
top_10
```

### Exercise: Proportions for popular names

Use the code block below to plot a line graph of `prop` vs `year` for each name in `top_10`. Be sure to color the lines by name to make the graph interpretable.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
top_10 |>
  ggplot() +
    geom_line(aes(x = year, y = prop, color = name))
```

:::

### Exercise: Total children for popular names

Now use `top_10` to plot `n` vs `year` for each of the names. How are the plots different? Why might that be? How does this affect our decision to use total children as a measure of popularity?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
top_10 |>
  ggplot() +
    geom_line(aes(x = year, y = n, color = name))
```

:::

###

Good job! This graph shows different trends than the one above, now let's consider why.

##

```{r}
#| echo: false
#| results: asis
create_buttons("04-mutate.html")
```
