---
title: "Deriving information with {dplyr}"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```

In this case study, you will identify the most popular American names from 1880 to 2015. While doing this, you will master three more dplyr functions:

* `mutate()`, `group_by()`, and `summarize()`, which help you use your data to compute new variables and summary statistics

These are some of the most useful R functions for data science, and this tutorial provides everything you need to learn them. 

This tutorial uses the [core tidyverse packages](http://tidyverse.org/), including {ggplot2}, {tibble}, and {dplyr}, as well as the `babynames` package. All of these packages have been pre-installed and pre-loaded for your convenience.


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-most-popular-names.html")
```
