---
title: "`arrange()`"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
filters:
  - webr
webr:
  packages:
    - babynames
    - dplyr
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(babynames)

source(here::here("R", "helpers.R"))
```

`arrange()` returns all of the rows of a data frame reordered by the values of a column. As with `select()`, the first argument of `arrange()` should be a data frame and the remaining arguments should be the names of columns. If you give `arrange()` a single column name, it will return the rows of the data frame reordered so that the row with the lowest value in that column appears first, the row with the second lowest value appears second, and so on. If the column contains character strings, `arrange()` will place them in alphabetical order.

### Exercise: `arrange()`

Use the code chunk below to arrange babynames by `n`. Can you tell what the smallest value of `n` is? 

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
arrange(babynames, n)
```

:::

###

Good job! The compiler of `babynames` used 5 as a cutoff; a name only made it into `babynames` for a given year and gender if it was used for five or more children.

### Tie breakers

If you supply additional column names, `arrange()` will use them as tie breakers to order rows that have identical values in the earlier columns. Add to the code below, to make `prop` a tie breaker. The result should first order rows by value of `n` and then reorder rows within each value of `n` by values of `prop`. 

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
arrange(babynames, n)


```

## {{< fa circle-check >}}&ensp;Solution

```r
arrange(babynames, n, prop)
```

:::


### `desc()`

If you would rather arrange rows in the opposite order, i.e. from _large_ values to _small_ values, surround a column name with `desc()`. `arrange()` will reorder the rows based on the largest values to the smallest.

Add a `desc()` to the code below to display the most popular name for 2017 (the largest year in the dataset) instead of 1880 (the smallest year in the dataset).

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
arrange(babynames, year, desc(prop))


```

## {{< fa circle-check >}}&ensp;Solution

```r
arrange(babynames, desc(year), desc(prop))
```

:::

Think you have it? Click Continue to test yourself.

### `arrange()` quiz

Which name was the most popular for a single gender in a single year? In the code chunk below, use `arrange()` to make the row with the largest value of `prop` appear at the top of the data set.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
arrange(babynames, desc(prop))
```

:::

Now arrange `babynames` so that the row with the largest value of `n` appears at the top of the data frame. Will this be the same row? Why or why not?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
arrange(babynames, desc(n))
```

:::

###

The number of children represented by each proportion grew over time as the population grew.

##

```{r}
#| echo: false
#| results: asis
create_buttons("05-pipe.html")
```
