---
title: "Working with tibbles"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```

In this primer, you will explore the popularity of different names over time. To succeed, you will need to master some common tools for manipulating data with R:

* tibbles and `View()`, which let you inspect raw data
* `select()` and `filter()`, which let you extract rows and columns from a data frame
* `arrange()`, which lets you reorder the rows in your data
* `|>`, which organizes your code into reader-friendly "pipes"
* `mutate()`, `group_by()`, and `summarize()`, which help you use your data to compute new variables and summary statistics

These are some of the most useful R functions for data science, and the tutorials that follow will provide you everything you need to learn them. 

In the tutorials, we'll use a dataset named `babynames`, which comes in a package that is also named `babynames`. Within `babynames`, you will find information about almost every name given to children in the United States since 1880. 

This tutorial introduces `babynames` as well as a new data structure that makes working with data in R easy: the tibble.

In addition to `babynames`, this tutorial uses the [core tidyverse packages](http://tidyverse.org/), including {ggplot2}, {tibble}, and {dplyr.} All of these packages have been pre-installed for your convenience. But they haven't been pre-loaded---something you will soon learn more about!


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-babynames.html")
```
