---
title: "Exploratory data analysis"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

options(
  digits = 3,
  width = 300
)

library(tidyverse)
library(palmerpenguins)
library(checkdown)

source(here::here("R", "helpers.R"))
```


### What is EDA? {.no-hide}

EDA is an iterative cycle that helps you understand what your data says. When you do EDA, you:

1. Generate questions about your data

1. Search for answers by visualizing, transforming, and/or modeling your data

1. Use what you learn to refine your questions and/or generate new questions

EDA is an important part of any data analysis. You can use EDA to make discoveries about the world; or you can use EDA to ensure the quality of your data, asking questions about whether the data meets your standards or not.

### The EDA mindset

EDA is not a formal process with a strict set of rules. More than anything, EDA is a state of mind. During the initial phases of EDA, you should feel free to investigate every idea that occurs to you. Some of these ideas will pan out, and some will be dead ends. As your exploration continues, you will home in on lines of inquiry that reveal insights worth writing up and communicating to others.

### Questions

Your goal during EDA is to develop an understanding of your data. The easiest way to do this is to use questions as tools to guide your investigation. When you ask a question, the question focuses your attention on a specific part of your dataset and helps you decide which graphs, models, or transformations to make.

> "Far better an approximate answer to the right question, which is often
> vague, than an exact answer to the wrong question, which can always be made
> precise." 
> 
> --- John Tukey

### Quantity vs Quality

EDA is, fundamentally, a creative process. And like most creative processes, the key to asking _quality_ questions is to generate a large _quantity_ of questions. It is difficult to ask revealing questions at the start of your analysis because you do not know what insights are contained in your dataset. On the other hand, each new question that you ask will highlight a new aspect of your data and increase your chance of making a discovery. You can quickly drill down into the most interesting parts of your data---and develop a set of thought-provoking questions---if you follow up each question with a new question based on what you find.

> "There are no routine statistical questions, only questionable statistical
> routines." 
> 
> --- Sir David Cox

### Two useful questions

There is no rule about which questions you should ask to guide your research. However, two types of questions will always be useful for making discoveries within your data. You can loosely word these questions as:

1. What type of **variation** occurs **within** my variables?

1. What type of **covariation** occurs **between** my variables?

The rest of this tutorial will look at these two questions. To make the discussion easier, let's define some terms...

### Definitions

* A __variable__ is a quantity, quality, or property that you can measure. 

* A __value__ is the state of a variable when you measure it. The value of a variable may change from measurement to measurement.
  
* An __observation__ or __case__ is a set of measurements made under similar conditions (you usually make all of the measurements in an observation at the same time and on the same object). An observation will contain several values, each associated with a different variable. I'll sometimes refer to an observation as a case or data point.

* __Tabular data__ is a table of values, each associated with a variable and an observation. Tabular data is __tidy__ if each value is placed in its own cell, each variable in its own column, and each observation in its own row. 
    
  So far, all of the data that you've seen has been tidy. In real-life, most data isn't tidy, so we'll come back to these ideas again in [Data Wrangling](/tidy-data/01-reshape-data/index.qmd).


### Review 1: Discovery or Confirmation?

You can think of science as a process with two steps: _discovery_ and _confirmation_. Scientists first observe the world to discover a hypothesis to test. Then, they devise a test to confirm the hypotheses against new data. If a hypothesis survives many tests, scientists begin to trust that it is a reliable explanation of the data.

The separation between discovery and confirmation is especially important for data scientists. It is easy for patterns to appear in data by coincidence. As a result, data scientists first look for patterns, and then try to confirm that the patterns exist in the real world. Sometimes this confirmation requires computing the probability that the pattern is due to random chance, a task that often involves collecting new data.

::: {.callout-note appearance="simple" icon=false .question}

**Is EDA a tool for discovery or confirmation?**

```{r echo=FALSE}
check_question(
  answer = "Discovery", 
  options = c(
    "Discovery",
    "Confirmation"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct! EDA is a tool for discovery; in fact, EDA is one of the most fruitful tools for discovery in science. We'll focus on discovery throughout this primer, but remember that you should test any pattern that you discover before you rely on it."),
  wrong = "Not quite. EDA reveals patterns, but it does not confirm that those patterns exist beyond your data set."
)
```
:::


### Review 2: Quality or Quantity?

::: {.callout-note appearance="simple" icon=false .question}

**When you begin to explore data, is it better to formulate one or two high-quality questions to ask, or many, many questions to explore?**

```{r echo=FALSE}
check_question(
  answer = "Many, many questions", 
  options = c(
    "One or two high-quality questions",
    "Many, many questions"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 25,
  right = c("Correct! Each question you ask creates a new opportunity to discover something surprising. You can lead yourself to high-value questions by iterating on questions that reveal unexpected results."),
  wrong = "Not quite. It is hard to spot useful questions ahead of time because you don't know which discoveries are waiting in your data."
)
```
:::


### Review 3: Definitions

`penguins` is a fun example dataset that comes with [the {palmerpenguins} package](https://allisonhorst.github.io/palmerpenguins/). The data set describes 344 penguins spread across 3 islands in the Palmer Archipelago, Antarctica. Each row in `penguins` displays a details about a bird, including the island it lives on, its bill length, its bill depth (i.e. how tall the bill is), its flipper length, its body mass, its sex, and the year its measurements were taken. You can use these measurements to deduce the penguin's species, which is also included in `penguins`.

```{r}
penguins
```

::: {.callout-note appearance="simple" icon=false .question}

**Which of these is a *variable* in the `penguins` dataset?**

```{r echo=FALSE}
check_question(
  answer = "bill_length_mm", 
  options = c(
    "Torgerson",
    "bill_length_mm",
    "bird",
    "18.1"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 2,
  right = c("Correct!")
)
```

:::

::: {.callout-note appearance="simple" icon=false .question}
**Which of these is a *value* in the `penguins` dataset?**

```{r echo=FALSE}
check_question(
  answer = "39.1", 
  options = c(
    "species",
    "body_mass_g",
    "39.1",
    "bird"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 3,
  right = c("Correct!")
)
```

:::

::: {.callout-note appearance="simple" icon=false .question}
**Which of these is an *observation* in the `penguins` dataset?**

```{r echo=FALSE}
check_question(
  answer = "The collection of measurements, `Adelie`, `Torgerson`, `39.1`, `18.7`, `181`, `3750`, `male`, and `2007`, which describe the first penguin in the dataset.", 
  options = c(
    "The collection of measurements, `39.1`, `39.5`, `40.3`, and so on, which are all of the values in the `bill_length_mm` column.",
    "The collection of names, `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`.",
    "The collection of measurements, `Adelie`, `Torgerson`, `39.1`, `18.7`, `181`, `3750`, `male`, and `2007`, which describe the first penguin in the dataset."
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 4,
  right = c("Correct! These measurements were all collected under similar circumstances: on the same penguin, presumably at the same time. If a relationship exists between the variables that these values describe, we would expect the relationship to also exist between these values.")
)
```
:::


##

```{r}
#| echo: false
#| results: asis
create_buttons("02-variation.html")
```
