---
title: "Facets"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr
filters:
  - webr
webr:
  packages:
    - ggplot2
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```

### Facetting {.no-hide}

You can more easily compare subgroups of data if you place each subgroup in its own subplot, a process known as **facetting.**

```{r echo=FALSE}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap(vars(cut))
```

### `facet_grid()`

{ggplot2} provides two functions for facetting. `facet_grid()` divides the plot into a grid of subplots based on the values of one or two facetting variables. To use it, add `facet_grid()` to the end of your plot call. 

The code chunks below show three ways to facet with `facet_grid()`. Spot the differences between the chunks, then run the code to learn what the differences do.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(rows = vars(clarity), cols = vars(cut))


```

:::

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(cols = vars(cut))


```

:::

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(rows = vars(clarity))


```

:::



### `facet_grid()` recap

As you saw in the code examples, you use `facet_grid()` by passing a `rows` and/or a `cols` argument, with the names of the variables inside a `vars()` function.

* `facet_grid()` will split the plot into facets vertically by the values of the `rows` variable: each facet will contain the observations that have a common value of the variable. 
* `facet_grid()` will split the plot horizontally by values of the `cols` variable. The result is a grid of facets, where each specific subplot shows a specific combination of values.


### `facet_wrap()`

`facet_wrap()` provides a more relaxed way to facet a plot on a _single_ variable. It will split the plot into subplots and then reorganize the subplots into multiple rows so that each plot has a more or less square aspect ratio. In short, `facet_wrap()` _wraps_ the single row of subplots that you would get with `facet_grid()` into multiple rows. 

To use `facet_wrap()` pass it a variable name inside `vars()`, e.g. `facet_wrap(vars(color))`.

Add `facet_wrap()` to the code below to create the graph that appeared at the start of this section. Facet by `cut`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut))


```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap(vars(cut))
```

:::


### `scales`

By default, each facet in your plot will share the same $x$ and $y$ ranges. You can change this by adding a `scales` argument to `facet_wrap()` or `facet_grid()`.

* `scales = "free"` will let the $x$ and $y$ range of each facet vary
* `scales = "free_x"` will let the $x$ range of each facet vary, but not the $y$ range
* `scales = "free_y"` will let the $y$ range of each facet vary, but not the $x$ range. This is a convenient way to compare the shapes of different distributions

Try changing the `scales` argument from `free` to `free_x` to `free_y` to see how it works:

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap(vars(cut), scales = "free")


```

:::



### Recap

In this tutorial, you learned how to make bar charts; but much of what you learned applies to other types of charts as well. Here's what you should know:

* Bar charts are the basis for histograms, which means that you can interpret histograms in a similar way.
* Bars are not the only geom in {ggplot2} that use the fill aesthetic. You can use both fill and color aesthetics with any geom that has an "interior" region.
* You can use the same position adjustments with any {ggplot2} geom: `"identity"`, `"stack"`, `"dodge"`, `"fill"`, `"nudge"`, and `"jitter"` (we'll learn about `"nudge"` and `"jitter"` later). Each geom comes with its own sensible default.
* You can facet any {ggplot2} plot by adding `facet_grid()` or `facet_wrap()` to the plot call.

Bar charts are an excellent way to display the distribution of a categorical variable. In the next tutorial, we'll meet a set of geoms that display the distribution of a continuous variable.


##

```{r}
#| echo: false
#| results: asis
create_buttons(NULL)
```
