---
title: "Your name"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(babynames)

source(here::here("R", "helpers.R"))
```

### The history of your name {.no-hide}

You can use the data in `babynames` to make graphs like this, which reveal the history of a name, perhaps your name.

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="90%"}
babynames |> 
  filter(name == "Andrew", sex == "M") |> 
  ggplot() +
    geom_line(aes(x = year, y = prop)) +
    labs(title = "Popularity of the name Andrew")
```

But before you do, you will need to trim down `babynames`. At the moment, there are more rows in `babynames` than you need to build your plot. 

### An example

To see what I mean, consider how I made the plot above: I began with the entire dataset, which if plotted as a scatterplot would've looked like this.

```{r plot-all-names, out.width="60%", cache=TRUE}
ggplot(babynames) +
  geom_point(aes(x = year, y = prop)) +
  labs(title = "Popularity of every name in the dataset")
```

I then narrowed the data to just the rows that contain my name, before plotting the data with a line geom. Here's how the rows with just my name look as a scatterplot.

```{r out.width="60%"}
babynames |> 
  filter(name == "Andrew", sex == "M") |> 
  ggplot() +
    geom_point(aes(x = year, y = prop)) +
    labs(title = "Popularity of the name Andrew")
```

If I had skipped this step, my line graph would've connected all of the points in the large dataset, creating an uninformative graph.

```{r out.width="60%", cached=TRUE}
ggplot(babynames) +
  geom_line(aes(x = year, y = prop)) +
  labs(title = "Popularity of every name in the dataset")
```

Your goal in this section is to repeat this process for your own name (or a name that you choose). Along the way, you will learn a set of functions that isolate information within a dataset. 

### Isolating data

This type of task occurs often in data science: you need to extract data from a table before you can use it. You can do this task quickly with three functions that come in the {dplyr} package:

1. **`select()`**, which extracts columns from a data frame
1. **`filter()`**, which extracts rows from a data frame
1. **`arrange()`**, which moves important rows to the top of a data frame

Each function takes a data frame or tibble as its first argument and returns a new data frame or tibble as its output. 


##

```{r}
#| echo: false
#| results: asis
create_buttons("02-select.html")
```
