---
title: "`select()`"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr

webr:
  packages:
    - babynames
    - dplyr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```

`select()` extracts columns of a data frame and returns the columns as a new data frame. To use `select()`, pass it the name of a data frame to extract columns from, and then the names of the columns to extract. The column names do not need to appear in quotation marks or be prefixed with a `$`; `select()` knows to find them in the data frame that you supply.

### Exercise: `select()`

Use the example below to get a feel for `select()`. Can you extract just the `name` column? How about the `name` and `year` columns? How about all of the columns except `prop`?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
select(babynames, name, sex)


```

## {{< fa circle-check >}}&ensp;Solution

```r
select(babynames, name)
select(babynames, name, year)
select(babynames, year, sex, name, n)
```

:::


### `select()` helpers

You can also use a series of helpers with `select()`. For example, if you place a minus sign before a column name, `select()` will return every column but that column. Can you predict how the minus sign will work here?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
select(babynames, -c(n, prop))


```

:::

The table below summarizes the other `select()` helpers that are available in {dplyr}. Study it, and then click "Continue" to test your understanding.

| Helper function     | Use                                          | Example                                         |
|-------------------|------------------------|------------------------------|
| **`-`**             | Columns except                               | `select(babynames, -prop)`                      |
| **`:`**             | Columns between (inclusive)                  | `select(babynames, year:n)`                     |
| **`contains()`**    | Columns that contains a string               | `select(babynames, contains("n"))`              |
| **`ends_with()`**   | Columns that ends with a string              | `select(babynames, ends_with("n"))`             |
| **`matches()`**     | Columns that matches a regex                 | `select(babynames, matches("n"))`               |
| **`num_range()`**   | Columns with a numerical suffix in the range | Not applicable with `babynames`                 |
| **`one_of()`**      | Columns whose name appear in the given set   | `select(babynames, one_of(c("sex", "gender")))` |
| **`starts_with()`** | Columns that starts with a string            | `select(babynames, starts_with("n"))`           |

: {tbl-colwidths="[15, 35, 35]" .striped .hover .table-sm}


### `select()` quiz

::: {.callout-note appearance="simple" icon=false .question}

**Which of these is not a way to select the `name` and `n` columns together?**

```{r predict, echo=FALSE}
check_question(
  answer = 'select(babynames, ends_with("n"))', 
  options = c(
    "select(babynames, -c(year, sex, prop))",
    "select(babynames, name:n)",
    'select(babynames, starts_with("n"))',
    'select(babynames, ends_with("n"))'
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct!")
)
```
:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("03-filter.html")
```
