---
title: "A code template"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr
filters:
  - webr
webr:
  packages:
    - ggplot2
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(checkdown)
library(here)

source(here::here("R", "helpers.R"))
```

> "The simple graph has brought more information to the data analyst’s mind than any other device." 
> 
> --- John Tukey

### 

Let's begin with a question to explore.

::: {.callout-note appearance="simple" icon=false .question}

**What do you think: Do cars with bigger engines use more fuel than cars with smaller engines?**

```{r predict, echo=FALSE}
check_question(
  answer = c(
   "Cars with bigger engines use more fuel.",
   "Cars with bigger engines use less fuel."
  ), 
  options = c(
   "Cars with bigger engines use more fuel.",
   "Cars with bigger engines use less fuel."
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("It could go either way! In other words, there might be a positive or a negative relationship between engine size and fuel efficiency. Now let's test your hypothesis with data.")
)
```
:::


### `mpg`

You can test your hypothesis with the `mpg` dataset that comes in the `ggplot2` package. `mpg` contains observations collected on 38 models of cars by the US Environmental Protection Agency. 

To see the `mpg` data frame, type `mpg` in the code block below and click "Run Code".

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
mpg
```

:::

###

Good job!  We'll use interactive code chunks like this throughout these tutorials. Whenever you encounter one, you can click Run Code to run (or re-run) the code in the chunk. If there is a Solution tab, you can click it to see the answer.

###

Among the variables in `mpg` are:

1. `displ`, a car's engine size, in liters.
1. `hwy`, a car's fuel efficiency on the highway, in miles per gallon (mpg). A car with a low mpg consumes more fuel than a car with a high mpg when they travel the same distance.

Now let's use this data to make our first graph.

### A plot

The code below uses functions from the **ggplot2** package to plot the relationship between `displ` and `hwy`. 

To see the plot, click "Run Code."


::: {.callout-note appearance="simple" icon=false .question}

**In the code below we are mapping data columns to graphical features. Which columns are being mapped to which graphical features? (Choose more than one option.) **

```{r}
#| label: ggplot-mapping
#| echo: FALSE
check_question(
  answer = c(
   "Data column `displ` is mapped to the x-axis.",
   "Data column `hwy` is mapped to the y-axis."
  ), 
  options = c(
   "Data column `displ` is mapped to the x-axis.",
   "Data column `hwy` is mapped to the y-axis.",
   "Data column `hwy` is mapped to the x-axis.",
   "Data column `displ` is mapped to the y-axis."
  ), 
  type = "checkbox", 
  button_label = "Submit answer", 
  q_id = 12,
  right = c("That's correct! The variables `displ` and `hwy` are two columns in the data frame. The first is being mapped to the x-axis, and the second is being mapped to the y-axis.")
)
```
:::


```{webr-r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```



Can you spot the relationship?

### And the answer is...

The plot shows a negative relationship between engine size (`displ`) and fuel efficiency (`hwy`). Points that have a large value of `displ` have a small value of `hwy` and vice versa.

In other words, cars with big engines use _more_ fuel. If that was your hypothesis, you were right!

Now let's look at how we made the plot.

### `ggplot()`

Here's the code that we used to make the plot. Notice that it contains three functions: `ggplot()`, `geom_point()`, and `aes()`.

```{r eval=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

In R, a function is a name followed by a set of parentheses. Many functions require special information to do their jobs, and you write this information between the parentheses.

### `ggplot`

The first function, `ggplot()`, creates a coordinate system that you can add layers to. The first argument of `ggplot()` is the dataset to use in the graph. 

By itself, `ggplot(data = mpg)` creates an empty graph, which looks like this: try it!


```{webr-r}

ggplot(data = mpg)


```


You should see an empty canvas. To plot something on the canvas, we need to add layers!


### `geom_point()`

`geom_point()` adds a layer of points to the empty plot created by `ggplot()`. This gives us a scatterplot. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### `mapping = aes()`

`geom_point()` takes a `mapping` argument that defines which variables in your dataset are mapped to which axes in your graph. The `mapping` argument is always paired with the function `aes()`, which you use to gather together all of the mappings that you want to create. 

### 

Here, we want to map the `displ` variable to the x axis and the `hwy` variable to the y axis, so we add `x = displ` and `y = hwy` inside of `aes()` (and we separate them with a comma). 

Where will ggplot2 look for these mapped variables? In the data frame that we passed to the `data` argument, in this case, `mpg`.

### A graphing workflow

Our code follows the common workflow for making graphs with ggplot2. To make a graph, you: 

1. Start the graph with `ggplot()`
1. Add elements to the graph with a `geom_` function
1. Map the data columns to the graphical features with  `mapping = aes()` argument

### A graphing template

In fact, you can turn our code into a reusable template for making graphs. To make a graph, replace the bracketed sections in the code below with a data set, a `geom_` function, or a collection of mappings.

Give it a try! Replace the bracketed sections with `mpg`, `geom_boxplot`, and `x = class, y = hwy` to make a slightly different graph. Be sure to delete the `#` symbols before you run the code.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
# ggplot(data = <DATA>) + 
#  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = class, y = hwy))
```

:::

###

Good job! This plot uses boxplots to compare the fuel efficiencies of different types of cars. ggplot2 comes with many geom functions that each add a different type of layer to a plot. You'll learn more about boxplots and other geoms in the tutorials that follow.

### Common problems

As you start to run R code, you're likely to run into problems. Don't worry---it happens to everyone. I have been writing R code for years, and every day I still write code that doesn't work! 

Start by carefully comparing the code that you're running to the code in the examples. R is extremely picky, and a misplaced character can make all the difference. Make sure that every `(` is matched with a `)` and every `"` is paired with another `"`. Also pay attention to capitalization; R is case sensitive.

### `+` location

One common problem when creating ggplot2 graphics is to put the `+` in the wrong place: it has to come at the end of a line, not the start. In other words, make sure you haven't accidentally written code like this:

```r
ggplot(data = mpg) 
+ geom_point(mapping = aes(x = displ, y = hwy))
```

### Help

If you're still stuck, try the help. You can get help about any R function by running `?function_name` in a code chunk, e.g. `?geom_point`. Don't worry if the help doesn't seem that helpful --- instead skip down to the bottom of the help page and look for a code example that matches what you're trying to do.

If that doesn't help, carefully read the error message that appears when you run your (non-working) code. Sometimes the answer will be buried there! But when you're new to R, you might not yet know how to understand the error message. Another great tool is Google: try googling the error message, as it's likely someone else has had the same problem, and has gotten help online.

### Exercise 1

Run `ggplot(data = mpg)`. What do you see?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

:::



Good job! A ggplot that has no layers looks blank. To finish the graph, add a geom function.

### Exercise 2 

Make a scatterplot of `cty` vs `hwy`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa lightbulb >}}&ensp;Hint

**Hint:** Scatterplots are also called points plots and bubble plots. They use the point geom.

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy))
```

:::



Excellent work!

### Exercise 3  

What happens if you make a scatterplot of `class` vs `drv`. Try it. Why is the plot not useful?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```

:::
    
###

Nice job! `class` and `drv` are both categorical variables. As a result, points can only appear at certain values, where many points overlap each other. You have no idea how many points fall on top of each other at each location. Experiment with `geom_count()` to find a better solution.

##


```{r}
#| echo: false
#| results: asis
create_buttons("02-aesthetic-mappings.html")
```
