---
title: "Aesthetics"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr
filters:
  - webr
webr:
  packages:
    - ggplot2
    - dplyr
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```


### Aesthetics for bars {.no-hide}

`geom_bar()` and `geom_col()` can use several aesthetics:

* `alpha`
* `color`
* `fill`
* `linetype`
* `size`

One of these, `color`, creates the most surprising results. Predict what the code below will return and then run it.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))


```

:::

### `fill`

The `color` aesthetic controls the outline of each bar in your bar plot, which may not be what you want. To color the interior of each bar, use the `fill` aesthetic:

```{r echo=FALSE, out.width="100%"}
#| layout-ncol: 2
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut), linewidth = 1) + 
  labs(title = "color = cut")

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut)) + 
  labs(title = "fill = cut")
```

Use the code chunk below to experiment with fill, along with other `geom_bar()` aesthetics, like `alpha`, `linetype`, and `size`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))


```

:::


### Width

You can control the width of each bar in your bar chart with the `width` parameter. In the chunk below, set `width = 1`, then `width = 0.5`. Can you spot the difference?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut), width = 0.9)


```

:::

Notice that width is a _parameter_, not an aesthetic mapping. Hence, you should set width _outside_ of the `aes()` function.

### Exercise 5: Aesthetics

Create a colored bar chart of the `class` variable from the `mpg` data set, which comes with ggplot2. Map the interior color of each bar to `class`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill = class))
```

:::


##

```{r}
#| echo: false
#| results: asis
create_buttons("03-position-adjustments.html")
```
