---
title: "Bar charts"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```


This tutorial will show you how to make and enhance **bar charts** with the {ggplot2} package. You will learn how to:

* make and interpret bar charts
* customize bar charts with **aesthetics** and **parameters**
* use **position adjustments**
* use **facets** to create subplots

The tutorial is adapted from _R for Data Science_ by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at [shop.oreilly.com](http://shop.oreilly.com/product/0636920034407.do).

The tutorial uses the {ggplot2} and {dplyr} packages, which have been pre-loaded for your convenience.


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-bar-charts.html")
```
