---
title: "Similar geoms"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
filters:
  - webr
webr:
  packages:
    - ggplot2
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)

source(here::here("R", "helpers.R"))
```

### `geom_dotplot()` {.no-hide}

Boxplots provide a quick way to represent a distribution, but they leave behind a lot of information. {ggplot2} supplements boxplots with two geoms that show more information.

The first is `geom_dotplot()`. If you set the `binaxis` parameter of `geom_dotplot()` to `"y"`, `geom_dotplot()` behaves like `geom_boxplot()`, display a separate distribution for each group of data.

Here each group functions like a vertical histogram. Add the parameter `stackdir = "center"` then re-run the code. Can you interpret the results?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = class, y = hwy), binaxis = "y", 
               dotsize = 0.5, binwidth = 1)


```


## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = class, y = hwy), binaxis = "y", 
               dotsize = 0.5, binwidth = 1, stackdir = "center")
```

:::

###

Good job! When you set `stackdir = "center"`, `geom_dotplot()` arranges each row of dots symmetrically around the $x$ value. This layout will help you understand the next geom.

As in the histogram tutorial, it takes a lot of tweaking to make a dotplot look right. As a result, I tend to only use them when I want to make a point.


### `geom_violin()`

`geom_violin()` provides a second alternative to `geom_boxplot()`. A violin plot uses densities to draw a smoothed version of the centered dotplot you just made.

You can think of a violin plot as an outline drawn around the edges of a centered dotplot. Each "violin" spans the range of the data. The violin is thick where there are many values, and thin where there are few.

Convert the plot below from a boxplot to a violin plot. Note that violin plots do not use the parameters you saw for dotplots.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))


```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = mpg) +
  geom_violin(mapping = aes(x = class, y = hwy))
```

:::

###

'Good job! Another way to interpret a violin plot is to mentally "push" the width of each violin all to one side (so the other side is a straight line). The result would be a density (e.g. `geom_density()`) turned on its side for each distribution).

### Exercise 7: Violin plots

You can further enhance violin plots by adding the parameter `draw_quantiles = c(0.25, 0.5, 0.75)`. This will cause ggplot2 to draw horizontal lines across the violins at the 25th, 50th, and 75th percentiles. These are the same three horizontal lines that are displayed in a boxplot (the 25th and 75th percentiles are the bounds of the box, the 50th percentile is the median).

Add these lines to the violin plot below.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
ggplot(data = mpg) +
  geom_violin(mapping = aes(x = class, y = hwy))


```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(data = mpg) +
  geom_violin(mapping = aes(x = class, y = hwy), draw_quantiles = c(0.25, 0.5, 0.75))
```

:::

###

Good job! Can you predict how you would use `draw_quantiles` to draw a horizontal line at a different percentile, like the 60th percentile?.


##

```{r}
#| echo: false
#| results: asis
create_buttons("03-counts.html")
```
