---
title: "Scatterplots"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```

A **scatterplot** displays the relationship between two continuous variables. Scatterplots are one of the most common types of graphs---in fact, you've met scatterplots already in [Visualization Basics](/basics/01-visualization-basics/). 

In this tutorial, you'll learn how to:

* Make new types of scatterplots with `geom_text()` and `geom_jitter()`
* Add multiple **layers** of geoms to a plot
* Enhance scatterplots with `geom_smooth()`, `geom_rug()`, and `geom_repel()`
* Change the **coordinate system** of a plot

The tutorial is adapted from [_R for Data Science_](https://r4ds.had.co.nz/) by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at [shop.oreilly.com](http://shop.oreilly.com/product/0636920034407.do).

The tutorial uses the {ggplot2}, {ggrepel}, and {dplyr} packages, which have been pre-loaded for your convenience.


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-scatterplots.html")
```
