---
title: "Exploratory data analysis"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```


This tutorial will show you how to explore your data in a systematic way, a task that statisticians call **exploratory data analysis**, or **EDA** for short. In the tutorial you will:

* Learn a strategy for exploring data
* Practice finding patterns in data
* Get tips about how to use different types of plots to explore data

The tutorial is excerpted from _R for Data Science_ by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at [shop.oreilly.com](http://shop.oreilly.com/product/0636920034407.do).


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-eda.html")
```
