---
title: "Reshape data"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```

The tools that you learned in the previous Primers work best when your data is organized in a specific way. This format is known as **tidy data** and it appears throughout the tidyverse. You will spend a lot of time as a data scientist wrangling your data into a usable format, so it is important to learn how to do this fast. 

This tutorial will teach you how to recognize tidy data, as well as how to reshape untidy data into a tidy format. In it, you will learn the core data wrangling functions for the tidyverse:

* `pivot_longer()`, which reshapes wide data into long data, and 
* `pivot_wider()`, which reshapes long data into wide data

This tutorial uses the [core tidyverse packages](http://tidyverse.org/), including {ggplot2}, {dplyr}, and {tidyr}, as well as the {babynames} package. All of these packages have been pre-installed and pre-loaded for your convenience.


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-tidy-data.html")
```
