---
title: "Welcome"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```

Visualization is one of the most important tools for data science. 

It is also a great way to start learning R; when you visualize data, you get an immediate payoff that will keep you motivated as you learn. After all, learning a new language can be hard!

This tutorial will teach you how to visualize data with R's most popular visualization package, `ggplot2`.

### 

The tutorial focuses on three basic skills:

1. How to create graphs with a reusable **template**
1. How to add variables to a graph with **aesthetics**
1. How to make different "types" of graphs with **geoms** 

In this tutorial, we will use the [core tidyverse packages](http://tidyverse.org/), including `ggplot2`. I've already loaded the packages for you, so let's begin! 

***

These examples are excerpted from _R for Data Science_ by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at [shop.oreilly.com](http://shop.oreilly.com/product/0636920034407.do).


##

```{r}
#| echo: false
#| results: asis
create_buttons("02-code-template.html")
```
