---
title: "tidyverse"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr

webr:
  packages:
    - tidyverse
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```


### The tidyverse {.no-hide}

The {tibble} package is one of several packages that are known collectively as ["the tidyverse"](http://tidyverse.org). Tidyverse packages share a common philosophy and are designed to work well together. For example, in this tutorial you will use the {tibble} package, the {ggplot2} package, and the {dplyr} package, all of which belong to the tidyverse.

### The tidyverse package

When you use tidyverse packages, you can make your life easier by using the {tidyverse} package. The {tidyverse} package provides a shortcut for installing and loading the entire suite of packages in "the tidyverse", e.g. 

```{r eval = FALSE}
install.packages("tidyverse")
library(tidyverse)
```

### Installing the tidyverse

Think of the {tidyverse} package as a placeholder for the packages that are in the "tidyverse". By itself, {tidyverse} does not do much, but when you install the {tidyverse} package it instructs R to install every other package in the tidyverse at the same time. In other words, when you run `install.packages("tidyverse")`, R installs the following packages for you in one simple step:

* ggplot2
* dplyr
* tidyr
* readr
* purrr
* tibble
* hms
* stringr
* lubridate
* forcats
* DBI
* haven
* jsonlite
* readxl
* rvest
* xml2
* modelr
* broom

### Loading the tidyverse

When you load tidyverse with `library("tidyverse")`, it instructs R to load _the most commonly used_ tidyverse packages. These are:

* ggplot2
* dplyr
* tidyr
* readr
* purrr
* tibble
* stringr
* forcats
* lubridate

You can load the less commonly used tidyverse packages in the normal way, by running `library(<PACKAGE NAME>)` for each of them.

Let's give this a try. We will use the ggplot2 and dplyr packages later in this tutorial. Let's use the tidyverse package to load them in the chunk below:

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
library(tidyverse)
```

:::

### Quiz

::: {.callout-note appearance="simple" icon=false .question}

**Which package is not loaded by `library("tidyverse")`**

```{r tidyverse-check, echo=FALSE}
check_question(
  answer = "babynames", 
  options = c(
   "ggplot2",
   "dplyr",
   "tibble",
   "babynames"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct!")
)
```
:::

### Recap

Tibbles and the {tidyverse} package are two tools that make life with R easier. Ironically, you may not come to appreciate their value right away: these tutorials pre-load packages for you. However, you will want to use tibbles and the {tidyverse} package when you move out of the tutorials and begin doing your own work with R inside of RStudio.

This tutorial also introduced the babynames dataset. In the next tutorial, you will use this data set to plot the popularity of _your_ name over time. Along the way, you will learn how to filter and subset data sets in R.

##

```{r}
#| echo: false
#| results: asis
create_buttons(NULL)
```
