---
title: "The most popular names"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(babynames)
library(checkdown)

source(here::here("R", "helpers.R"))
```

### What are the most popular names of all time? {.no-hide}

Let's use `babynames` to answer a different question: what are the most popular names of all time?

This question seems simple enough, but to answer it we need to be more precise: how do you define "the most popular" names? Try to think of several definitions and then click Continue. After the Continue button, I will suggest two definitions of my own.

### Two definitions of popular

I suggest that we focus on two definitions of _popular_, one that uses sums and one that uses ranks:

1. **Sums** - A name is popular _if the total number of children that have the name is large when you sum across years_.
2. **Ranks** - A name is popular _if it consistently ranks among the top names from year to year_.

This raises a question:

::: {.callout-note appearance="simple" icon=false .question}

**Do we have enough information in `babynames` to compare the popularity of names?**

```{r predict, echo=FALSE}
check_question(
  answer = "Yes. We can use the information in `babynames` to compute the values we want.", 
  options = c(
   "No. No cell in `babynames` contains a rank value or a sum across years.",
   "Yes. We can use the information in `babynames` to compute the values we want."
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct!")
)
```
:::

### Deriving information

Every data frame that you meet implies more information than it displays. For example, `babynames` does not display the total number of children who had your name, but `babynames` certainly implies what that number is. To discover the number, you only need to do a calculation:

```{r}
babynames |> 
  filter(name == "Andrew", sex == "M") |> 
  summarize(total = sum(n))
```

### Useful functions

{dplyr} provides three functions that can help you reveal the information implied by your data: 

* `summarize()`  
* `group_by()`  
* `mutate()`

Like `select()`, `filter()` and `arrange()`, these functions all take a data frame as their first argument and return a new data frame as their output, which makes them easy to use in pipes.

Let's master each function and use them to analyze popularity as we go.

##

```{r}
#| echo: false
#| results: asis
create_buttons("02-summarize.html")
```
