---
title: "`summarize()`"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr

webr:
  packages:
    - babynames
    - dplyr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(babynames)

source(here::here("R", "helpers.R"))
```

`summarize()` takes a data frame and uses it to calculate a new data frame of summary statistics. 

### Syntax

To use `summarize()`, pass it a data frame and then one or more named arguments. Each named argument should be set to an R expression that generates a single value. Summarise will turn each named argument into a column in the new data frame. The name of each argument will become the column name, and the value returned by the argument will become the column contents.

Importantly, the `summarize()` function is *destructive*. It collapses a dataset into a single row and throws away any columns that we don’t use when summarizing. Watch this little animation to see what it does:

```{=html}
<video controls width="100%" style="display: block; margin: auto;">
  <source src="video/grp-summarize-00.mp4" type="video/mp4">
</video>
```

### Example

I used `summarize()` earlier to calculate the total number of boys named "Andrew", but let's expand that code to also calculate 

* `max`: the maximum number of boys named "Andrew" in a single year
* `mean`: the mean number of boys named "Andrew" per year

```{r}
babynames |> 
  filter(name == "Andrew", sex == "M") |> 
  summarize(total = sum(n), max = max(n), mean = mean(n))
```

Don't let the code above fool you. The first argument of `summarize()` is always a data frame, but when you use `summarize()` in a pipe, the first argument is provided by the pipe operator, `|>`. Here the first argument will be the data frame that is returned by `babynames |> filter(name == "Andrew", sex == "M")`.

### Exercise: `summarize()`

Use the code chunk below to compute three statistics:

1. the total number of children who ever had your name
1. the maximum number of children given your name in a single year
1. the mean number of children given your name per year 

If you cannot think of an R function that would compute each statistic, click the Solution tab.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
babynames |> 
  filter(name == "Andrew", sex == "M") |> 
  summarize(total = sum(n), max = max(n), mean = mean(n))
```

:::


### Summary functions

So far our `summarize()` examples have relied on `sum()`, `max()`, and `mean()`. But you can use any function in `summarize()` so long as it meets one criteria: the function must take a _vector_ of values as input and return a _single_ value as output. Functions that do this are known as **summary functions** and they are common in the field of descriptive statistics. Some of the most useful summary functions include:

1. **Measures of location**: `mean(x)`, `median(x)`, `quantile(x, 0.25)`, `min(x)`, and `max(x)`
1. **Measures of spread**: `sd(x)`, `var(x)`, `IQR(x)`, and `mad(x)`
1. **Measures of position**: `first(x)`, `nth(x, 2)`, and `last(x)`
1. **Counts**: `n_distinct(x)` and `n()`, which takes no arguments, and returns the size of the current group or data frame. 
1. **Counts and proportions of logical values**: `sum(!is.na(x))`, which counts the number of `TRUE`s returned by a logical test; `mean(y == 0)`, which returns the proportion of `TRUE`s returned by a logical test.

Let's apply some of these summary functions. Click Continue to test your understanding. 

### Khaleesi challenge

"Khaleesi" is a very modern name that appears to be based on the _Game of Thrones_ TV series, which premiered on April 17, 2011. In the chunk below, filter `babynames` to just the rows where `name == "Khaleesi"`. Then use `summarize()` and a summary function to return the first value of `year` in the data set.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
babynames |> 
  filter(name == "Khaleesi") |> 
  summarize(year = first(year))
```

:::


### Distinct name challenge

In the chunk below, use `summarize()` and a summary function to return a data frame with two columns:

* A column named `n` that displays the total number of rows in `babynames`
* A column named `distinct` that displays the number of distinct names in `babynames`

Will these numbers be different? Why or why not?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
babynames |> 
  summarize(n = n(), distinct = n_distinct(name))
```

:::

###

Good job! The two numbers are different because most names appear in the data set more than once. They appear once for each year in which they were used.

### `summarize()` by groups?

How can we apply `summarize()` to find the most popular names in `babynames`? You've seen how to calculate the total number of children that have your name, which provides one of our measures of popularity, i.e. the total number of children that have a name:

```{r eval=FALSE}
babynames |> 
  filter(name == "Andrew", sex == "M") |> 
  summarize(total = sum(n))
```

However, we had to isolate your name from the rest of your data to calculate this number. You could imagine writing a program that goes through each name one at a time and: 

1. filters out the rows with just that name
2. applies summarize to the rows

Eventually, the program could combine all of the results back into a single data set. However, you don't need to write such a program; this is the job of {dplyr}'s `group_by()` function.

##

```{r}
#| echo: false
#| results: asis
create_buttons("03-group_by.html")
```
