---
title: "Similar geoms"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr

webr:
  packages:
    - ggplot2
    - dplyr
    - forcats
    - gapminder
    - maps
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

library(tidyverse)
library(gapminder)
library(checkdown)

asia <- 
  gapminder |> 
  filter(country %in% c("China", "Korea, Dem. Rep.", "Korea, Rep.", "Japan")) |> 
  mutate(country = fct_recode(country, "North Korea" = "Korea, Dem. Rep.", "South Korea" = "Korea, Rep."))

tx <- map_data("state", region = "texas")

source(here::here("R", "helpers.R"))
```

```{webr}
#| label: setup
#| include: false
asia <- 
  gapminder |> 
  filter(country %in% c("China", "Korea, Dem. Rep.", "Korea, Rep.", "Japan")) |> 
  mutate(country = fct_recode(country, "North Korea" = "Korea, Dem. Rep.", "South Korea" = "Korea, Rep."))

tx <- map_data("state", region = "texas")
```

### `geom_step()` {.no-hide}

`geom_step()` draws a line chart in a stepwise fashion. To see what I mean, change the geom in the plot below and rerun the code.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
ggplot(asia) +
  geom_line(mapping = aes(x = year, y = lifeExp, color = country, linetype = country))


```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(asia) +
  geom_step(mapping = aes(x = year, y = lifeExp, color = country, linetype = country))
```

:::

###

Good job! You can control whether the steps move horizontally first and then vertically or vertically first and then horizontally with the parameters `direction = "hv"` (the default) or `direction = "vh"`.

### `geom_area()`

`geom_area()` is similar to a line graph, but it fills in the area under the line. To see `geom_area()` in action, change the geom in the plot below and rerun the code.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
ggplot(economics) +
  geom_line(mapping = aes(x = date, y = unemploy))


```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(economics) +
  geom_area(mapping = aes(x = date, y = unemploy))
```

:::

###

Good job! Filling the space under the line gives you a new way to customize your plot.


### Review 2 - Set vs. map

Do you recall from [Visualization Basics](/basics/01-visualization-basics/) how you would set the fill of our plot to blue (instead of, say, map the `fill` to a variable)? Give it a try.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
ggplot(economics) +
  geom_area(mapping = aes(x = date, y = unemploy))


```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(economics) +
  geom_area(mapping = aes(x = date, y = unemploy), fill = "blue")
```

:::

###

Good job! Remember that you map aesthetics to variables inside of `aes()`. You set aesthetics to values outside of `aes()`.


### Accumulation

`geom_area()` is a great choice if your measurements represent the accumulation of objects (like unemployed people). Notice that the $y$ axis `geom_area()` always begins or ends at zero.

Perhaps because of this, `geom_area()` can be quirky when you have multiple groups. Run the code below. Can you tell what happens here?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
ggplot(asia) +
  geom_area(mapping = aes(x = year, y = lifeExp, fill = country))


```

:::


### Review 3: Position adjustments

If you answered that people in China were living to be 300 years old, you guessed wrong. 

`geom_area()` is stacking each group above the group below. As a result, the line that should display the life expectancy for China displays the combined life expectancy for all countries.

You can fix this by changing the position adjustment for `geom_area()`. Give it a try below. Change the position parameter from "stack" (the implied default) to `"identity"`. See [Bar Charts](/visualize-data/02-bar-charts/) if you'd like to learn more about position adjustments.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
ggplot(asia) +
  geom_area(mapping = aes(x = year, y = lifeExp, fill = country), alpha = 0.3)


```

## {{< fa circle-check >}}&ensp;Solution

```r
ggplot(asia) +
  geom_area(mapping = aes(x = year, y = lifeExp, fill = country), position = "identity", alpha = 0.3)
```

:::

###

Good job! You can further customize your graph by switching from `geom_area()` to `geom_ribbon()`. `geom_ribbon()` lets you map the bottom of the filled area to a variable, as well as the top. See `?geom_ribbon` if you'd like to learn more.


### `geom_path()`

`geom_line()` comes with a strange bedfellow, `geom_path()`. `geom_path()` draws a line between points like `geom_line()`, but instead of connecting points in the order that they appear along the $x$ axis, `geom_path()` connects the points in the order that they appear _in the data set_.

It starts with the observation in row one of the data and connects it to the observation in row two, which it then connects to the observation in row three, and so on.

### `geom_path()` example

To see how `geom_path()` does this, let's rearrange the rows in the `economics` dataset. We can reorder them by `unemploy` value. Now the data set will begin with the observation that had the lowest value of `unemploy`.

```{r}
economics2 <- economics |> 
  arrange(unemploy)
economics2
```

### `geom_path()` example continued

If we plot the reordered data with both `geom_line()` and `geom_path()` we get two very different graphs. 

```{r out.width="100%", fig.show='hold'}
#| layout-ncol: 2
ggplot(economics2) +
  geom_line(mapping = aes(x = date, y = unemploy))

ggplot(economics2) +
  geom_path(mapping = aes(x = date, y = unemploy))
```

The plot on the left uses `geom_line()`, hence the points are connected in order along the $x$ axis. The plot on the right uses `geom_path()`. These points are connected in the order that they appear in the dataset, which happens to put them in order along the $y$ axis.

### A use case

Why would you want to use `geom_path()`? The code below illustrates one particularly useful case. The `tx` dataset contains latitude and longitude coordinates saved in a specific order.

```{r eval=FALSE}
tx
```

```{r echo=FALSE}
as_tibble(tx)
```


### tx

What do you think happens when you plot the data in `tx`? Run the code to find out.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
ggplot(tx) +
  geom_path(mapping = aes(x = long, y = lat))


```

:::

###

Good job! `geom_path()` reveals how you can use what is essentially a line plot to make a map (this is a map of the state of Texas). There are other better ways to make maps in R, but this low tech method is surprisingly versatile.

### `geom_polygon()`

`geom_polygon()` extends `geom_path()` one step further: it connects the last point to the first and then colors the interior region with a fill. The result is a polygon.

```{r out.width="80%"}
ggplot(tx) +
  geom_polygon(mapping = aes(x = long, y = lat))
```

### Exercise 2: Shattered glass

What do you think went wrong in the plot of Texas below?

```{r out.width="80%", echo=FALSE}
set.seed(100)
rows <- c(rep(c(1:10), 100) + rep(sample(0:99 * 10), each = 10), 1001:1088)

tx[rows, ] |> 
  ggplot() +
    geom_polygon(mapping = aes(x = long, y = lat))
```

::: {.callout-note appearance="simple" icon=false .question}

**What went wrong?**

```{r echo=FALSE}
check_question(
  answer = "The rows in the dataset became out of order.", 
  options = c(
    "The rows in the dataset became out of order.",
    "The programmer did not set a fill aesthetic.",
    "The programmer used a line plot instead of a polygon plot."
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  right = c("Correct! It looks like someone messed with tx. tx and datasets like it will have an order variable that you can use to ensure that the data is in the correct order before you plot it.")
)
```
:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("03-maps.html")
```
