---
title: "Line plots"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
---

```{r include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "70%"
)

source(here::here("R", "helpers.R"))
```

A **line graph** displays a functional relationship between two continuous variables. A **map**  displays spatial data. The two may seem different, but they are made in similar ways. This tutorial will examine them both.

In this tutorial, you'll learn how to:

* Make new types of line plots with `geom_step()`, `geom_area()`, `geom_path()`, and `geom_polygon()`
* Avoid "whipsawing" with the group aesthetic
* Find and plot map data with `geom_sf()`
* Transform a coordinate system into a map projection with `coord_sf()`

The tutorial is adapted from _R for Data Science_ by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at [shop.oreilly.com](http://shop.oreilly.com/product/0636920034407.do).

The tutorial uses the {ggplot2}, {sf}, and {dplyr} packages, which have been pre-loaded for your convenience.


##

```{r}
#| echo: false
#| results: asis
create_buttons("01-line-graphs.html")
```
