---
title: "Functions"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
source(here::here("R", "helpers.R"))
```

### Functions {.no-hide}

Watch [this video](https://vimeo.com/220490105):

```{=html}
<div class="ratio ratio-4x3">
<iframe src="https://player.vimeo.com/video/220490105" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>
</div>
```

### Run a function

Can you use the `sqrt()` function in the chunk below to compute the square root of 962?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
sqrt(962)
```

:::

### Code

Use the code chunk below to examine the code that `sqrt()` runs.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
sqrt
```

:::

###

Good job! `sqrt` immediately triggers a low level algorithm optimized for performance, so there is not much code to see.

### lm

Compare the code in `sqrt()` to the code in another R function, `lm()`. Examine `lm()`'s code body in the chunk below.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
lm
```

:::


### Help pages

Wow! `lm()` runs a lot of code. What does it do? Open the help page for `lm()` in the chunk below and find out.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
?lm


```

## {{< fa circle-check >}}&ensp;Solution

```r
?lm
```

:::

###

Good job! `lm()` is R's function for fitting basic linear models. No wonder it runs so much code.


### Code comments

What do you think the chunk below will return? Run it and see. The result should be nothing. R will not run anything on a line after a `#` symbol. This is useful because it lets you write human readable comments in your code: just place the comments after a `#`. Now delete the `#` and re-run the chunk. You should see a result.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
# sqrt(962)


```

## {{< fa circle-check >}}&ensp;Solution

```r
sqrt(962)
```

:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("02-arguments.html")
```
