---
title: "Arguments"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```

### Arguments {.no-hide}

Watch [this video](https://vimeo.com/220490157):

```{=html}
<div class="ratio ratio-4x3">
<iframe src="https://player.vimeo.com/video/220490157" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>
</div>
```

### `args()`

`rnorm()` is a function that generates random variables from a normal distribution. Find the arguments of `rnorm()` using the `args()` function.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
args(rnorm)
```

:::

###

Good job! `n` specifies the number of random normal variables to generate. `mean` and `sd` describe the distribution to generate the random values with.

### Optional arguments

::: {.callout-note appearance="simple" icon=false .question}

**Which arguments of `rnorm()` are not optional?**

```{r predict, echo=FALSE}
check_question(
  answer = "n", 
  options = c(
   "n",
   "mean",
   "sd"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct! `n` is not an optional argument because it does not have a default value.")
)
```
:::

### `rnorm()` 1

Use `rnrom()` to generate 100 random normal values with a mean of 100 and a standard deviation of 15.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
rnorm(100, mean = 100, sd = 15)
```

:::

### `rnorm()` 2

Can you spot the error in the code below? Fix the code and then re-run it.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
rnorm(100, mu = 100, sd = 15)


```

## {{< fa lightbulb >}}&ensp;Hint

**Hint:** In math, $\mu$ (mu, pronounced "mew" or "moo") is a Greek letter that stands for the mean of a distribution.

## {{< fa circle-check >}}&ensp;Solution

```r
rnorm(100, mean = 100, sd = 15)
```

:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("03-objects.html")
```
