---
title: "Objects"
format:
  html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
filters:
  - webr
webr:
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

```{r include=FALSE}
source(here::here("R", "helpers.R"))
```

### Objects {.no-hide}

Watch [this video](https://vimeo.com/220493412):

```{=html}
<div class="ratio ratio-4x3">
<iframe src="https://player.vimeo.com/video/220493412" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>
</div>
```

### Object names

You can choose almost any name you like for an object, as long as the name does not begin with a number or a special character like `+`, `-`, `*`, `/`, `^`, `!`, `@`, or `&`. 

For instance, check out this list of some possible object names. Some are okay to use; some are invalid:

- `today`: This is fine
- `1st`: This is **bad**; it starts with a number
- `+1`: This is **bad**; it starts with a special character
- `vars`: This is fine
- `\^_^`: This is **bad**; it starts with a special character
- `foo`: This is fine


### Using objects

In the code chunk below, save the results of `rnorm(100, mean = 100, sd = 15)` to an object named `data`. Then, on a new line, call the `hist()` function on `data` to plot a histogram of the random values.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
data <- rnorm(100, mean = 100, sd = 15)
hist(data)
```

:::

### What if?

What do you think would happen if you assigned `data` to a new object named `copy`, like this? Run the code and then inspect both `data` and `copy`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
data <- rnorm(100, mean = 100, sd = 15)
copy <- data


```

## {{< fa circle-check >}}&ensp;Solution

```r
data <- rnorm(100, mean = 100, sd = 15)
copy <- data
data
copy
```

:::

###

Good job! R saves a copy of the contents in data to copy.

### Datasets

Objects provide an easy way to store datasets in R. In fact, R comes with many toy datasets pre-loaded. Examine the contents of `mtcars` to see a classic toy dataset. Hint: how could you learn more about the `mtcars` object?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}



```

## {{< fa circle-check >}}&ensp;Solution

```r
mtcars
```

:::

###

Good job! You can learn more about mtcars by examining its help page with `?mtcars`.

### `rm()`

What if you accidentally overwrite an object? If that object came with R or one of its packages, you can restore the original version of the object by removing your version with `rm()`. Run `rm()` on `mtcars` below to restore the mtcars data set.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr-r}
mtcars <- 1
mtcars


```

## {{< fa circle-check >}}&ensp;Solution

```r
mtcars <- 1
mtcars
rm(mtcars)
mtcars
```

:::

###

Good job! Unfortunately, `rm()` cannot help you if you overwrite one of your own objects.

##

```{r}
#| echo: false
#| results: asis
create_buttons("04-vectors.html")
```
