---
title: "Packages"
format:
  live-html: 
    theme: 
      - zephyr
      - ../../html/custom.scss
    knitr:
      opts_chunk:
        dev: "ragg_png"
        dpi: 300 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr

webr:
  packages:
    - tidyverse
---

{{< include ../../_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```

### Packages {.no-hide}

Watch [this video](https://vimeo.com/220490447):

```{=html}
<div class="ratio ratio-4x3">
<iframe src="https://player.vimeo.com/video/220490447" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>
</div>
```

### A common error

::: {.callout-note appearance="simple" icon=false .question}

**What does this common error message suggest? `object _____ does not exist.`**

```{r packages1, echo=FALSE}
check_question(
  answer = c("Either"), 
  options = c(
   "You misspelled your object name",
   "You've forgot to load the package that ____ comes in",
   "Either"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct!")
)
```

:::


### Load a package

In the code chunk below, load the {tidyverse} package. Whenever you load a package R will also load all of the packages that the first package depends on. {tidyverse} takes advantage of this to create a shortcut for loading several common packages at once. Whenever you load {tidyverse}, {tidyverse} also loads {ggplot2}, {dplyr}, {tibble}, {tidyr}, {readr}, {purrr}, {forcats}, {stringr}, and {lubridate}.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
library(tidyverse)
```

:::

###

Good job! R will keep the packages loaded until you close your R session. When you re-open R, you'll need to reload your packages.


### Quotes

Did you know `library()` is a special function in R? You can pass `library()` a package name in quotes, like `library("tidyverse")`, or not in quotes, like `library(tidyverse)`---both will work! That's often not the case with R functions.

In general, you should always use quotes unless you are writing the _name_ of something that is already loaded into R's memory, like a function, vector, or data frame.

### Install packages

But what if the package that you want to load is not installed on your computer? How would you install the {dplyr} package on your own computer?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
install.packages("dplyr")
```

:::

###

Good job! You only need to install a package once, unless you wish to update your local copying by reinstalling the package. Notice that `install.packages()` _always_ requires quotes around the package name.


### Congratulations!

Congratulations. You now have a formal sense for how the basics of R work. Although you may think of your self as a data scientist, this brief computer science background will help you as you analyze data. Whenever R does something unexpected, you can apply your knowledge of how R works to figure out what went wrong.


##

```{r}
#| echo: false
#| results: asis
create_buttons(NULL)
```
